---
title: "Earthquake Sample Model"
author: "Nishant"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Sample Data Collection Details
The sample data is taken by randomly selecting the region of south east asia covering the countries `Malaysia`, `Indomesia` etc..
The sample data taken over a period of one year from **10 Feb 2017** to **10 Feb 2018**.
The data model used here may or may not be used in final model as it all depends on the structure and 



```{r cars, message=FALSE, warning=FALSE, echo=FALSE}
library(caret)
library(ggplot2)
library(dplyr)
library(lubridate)
library(PerformanceAnalytics)

```

## Data Overview

We will overview our data and remove unwanted or low variance features.

```{r pressure, echo=FALSE}
#importing data

DataEq = read.csv("oneyear.csv", stringsAsFactors = FALSE)

summary(DataEq)

#removing NA columns

DataEq = DataEq[ , -7]

class(DataEq$time)

#converting time to usable format

DataEq$time = ymd_hms(as_datetime(DataEq$time))
DataEq$updated = ymd_hms(as_datetime(DataEq$updated))
DataEq$year = year(DataEq$time)
DataEq$month = month(DataEq$time)
DataEq$day = day(DataEq$time)

#pattern in width and magnitude of the earthquake

ggplot(DataEq, aes(x = mag, y = depth, col = magType))+
  geom_point(alpha = 0.6)+
  facet_wrap(~ month)+
  theme_light()

# Removing no variance data

eqClean = DataEq[ , c(-10, -11, -14, -19, -20,-21)]

```

## Correlation of numeric Data

We observe the correlation of numeric data.

```{r, message=FALSE, warning=FALSE}

numClean = eqClean[, c(-1, -6, -10, -11)]

chart.Correlation(numClean)

```


